A Plot is Worth a Thousand Tests: Assessing Residual Diagnostics with the Lineup Protocol

### 2023-05-01 paper.tex
use google ref for all citations

In addition, scatterplots of residuals plotted against the fitted values
and each of the explanatory variables, can be used to scrutinize their
relationships. If there are any visually discoverable patterns, the
model is potentially inadequate or incorrectly specified. In general,
one looks for noticeable departures from the model such as non-linear
pattern or heteroskedasticity. A non-linear pattern would suggest that
the model needs to have some additional non-linear terms.
Heteroskedasticity suggests that the error is dependent on the
predictors, and hence violates the independence assumption. However,
correctly judging whether NO pattern exists in a residual plot is a
difficult task for humans. Humans will almost always will see a pattern
(see \citet{kahneman}), so the question that really needs answering is
whether any pattern perceived is consistent with randomness, purely
sampling variability, or noise. It is especially difficult to teach this
to new analysts and students \citep{loy2021bringing}. To address this,
statistical tests are available to check for non-linear patterns
\citep[e.g.][]{ramsey_tests_1969}, and heteroskedasticity
\citep[e.g.][]{breusch_simple_1979}.

Linear regression is a well-established procedure, and there is
considerable literature describing diagnostic procedures, e.g.
\citet{draper1998applied}, \citet{montgomery1982introduction},
\citet{belsley_regression_1980}, \citet{cook_applied_1999} and
\citet{cook1982residuals}. Interestingly, despite the abundance of
conventional tests, ALL of these writings advise that plotting residuals
is an essential tool for diagnosing regression model problems:

@@

Graphical summaries where residuals are plotted against fitted values,
or other functions of the predictors (expected to be approximately
orthogonal to the residuals) are considered to be the most important
residual plots by \citet{cook_applied_1999}. Figure
\ref{fig:residual-plot-common-departures}A shows an example of an ideal
residual plot where points are symmetrically distributed around the
horizontal zero line (red), with no discernible patterns. There can be
various types of departures from this ideal pattern. Non-linearity,
heteroskedasticity and non-normality, shown in Figure
\ref{fig:residual-plot-common-departures}B,C,D are three commonly
checked departures.

@@

Figure \ref{fig:residual-plot-common-departures}D shows a scatterplot
where the residuals have a skewed distribution, as seen by the uneven
vertical spread. Unlike non-linearity and heteroskedasticity,
non-normality is usually detected with a different type of residual
plot: a histogram or normal probability plot. Because we focus on
scatterplots, non-normality is not one of the departures examined in
depth in this paper. (loy\_variations\_2016 discuss related work on
non-normality checking.)

@@

Many different hypothesis tests are available to detect specific model
defects. For example, the presence of heteroskedasticity can usually be
tested by applying the White test
\citep{white_heteroskedasticity-consistent_1980} or the Breusch-Pagan
(BP) test \citep{breusch_simple_1979}, which are both derived from the
Lagrange multiplier test \citep{silvey1959lagrangian} principle that
relies on the asymptotic properties of the null distribution. To test
specific forms of non-linearity, one may apply the F-test as a model
structural test to examine the significance of specific polynomial and
non-linear forms of the predictors, or the significance of proxy
variables as in the Ramsey Regression Equation Specification Error Test
(RESET) \citep{ramsey_tests_1969}. The Shapiro-Wilk (SW) normality test
\citep{shapiro1965analysis} is the most widely used test of
non-normality included by many of the statistical software programs. The
Jarque-Bera test \citep{jarque1980efficient} is also used to directly
check whether the sample skewness and kurtosis match a normal
distribution.

Table \ref{tab:example-residual-plot-table} displays the \(p\)-values
from the RESET, BP and SW tests applied to the residual plots in Figure
\ref{fig:residual-plot-common-departures}. The RESET test and BP test
were computed using the \texttt{resettest} and \texttt{bptest} functions
from the R package \texttt{lmtest}, respectively. The SW test was
computed using the \texttt{shapiro.test} from the core R package
\texttt{stats}. (Although we didn't use it, it's useful to know that the
R package \texttt{skedastic} \citep{skedastic} also contains a large
collection of functions to test for heteroskedasticity.) Although, the
RESET test is exact, it requires the selection of a power parameter.
\citet{ramsey_tests_1969} recommends a power of four, which we adopted
in our analysis. The BP and SW tests are approximate.

We would expect the RESET test for non-linearity to reject residual plot
B, the BP test for heteroskedasticity to reject the residual plot C, and
the SW test for non-normality to reject residual plot D, which they all
do and all tests also correctly fail to reject residual plot A.
Interestingly, the BP and SW tests also reject the residual plots
exhibiting structure that they weren't designed for.
\citet{cook1982residuals} explain that most residual-based tests for
particular types of departure from model assumptions are also sensitive
to other types of departures. This could be considered a Type III error
(XXXREF), where the null hypothesis of good residuals is correctly
rejected but for the wrong reason. Also, some types of departure can
have elements of other types of departure, for example, non-linearity
could be viewed as heteroskedasticity. Additionally, other data problems
such as outliers can trigger rejection (or not) of the null hypothesis
\citep{cook_applied_1999}.

@@

The examination of data plots to infer signal or patterns (or lack
thereof) is fraught with variation in the human ability to interpret and
decode the information embedded in a graph
\citep{cleveland_graphical_1984}. Human examination of diagnostic plots
can feel subjective. Ideally we would expect visual evaluation to be
done objectively with small amounts of human-to-human variation in
reading data plots, but we cannot assume this is the case.

In practice, over-interpretation of a single plot is common. For
instance, \citet{roy_chowdhury_using_2015} described a published example
where authors over-interpreted separation between gene groups from a
two-dimensional projection of a linear discriminant analysis.
\citet{roy_chowdhury_using_2015} showed that there were no differences
in the expression levels between the gene groups. One solution to
over-interpretation is to examine the plot in the context of natural
sampling variability assumed by the model, called the lineup protocol,
as proposed in \citet{buja_statistical_2009}.
\citet{majumder_validation_2013} showed that the lineup protocol is
analogous to the null hypothesis significance testing framework. The
protocol consists of \(m\) randomly placed plots, where one plot is the
data plot, and the remaining \(m - 1\) plots, referred to as the
\emph{null plots}, are constructed using the same graphical procedure as
the data plot but the data is replaced with null data that is generated
in a manner consistent with the null hypothesis, \(H_0\). Then, an
observer who has not seen the data plot is asked to point out the most
different plot from the lineup. Under \(H_0\), it is expected that the
data plot would have no distinguishable difference from the null plots,
and the probability that the observer correctly picks the data plot is
\(1/m\). If one rejects \(H_0\) as the observer correctly picks the data
plot, then the Type I error of this test is \(1/m\). This protocol
requires a priori specification of \(H_0\) (or at least a null data
generating mechanism), much like the requirement of knowing the sampling
distribution of the test statistic in null hypothesis significance
testing framework.

@@

Data used in the \(m - 1\) null plots needs to be simulated. In
regression diagnostics, sampling data consistent with \(H_0\) is
equivalent to sampling data from the assumed model. As
\citet{buja_statistical_2009} suggested, \(H_0\) is usually a composite
hypothesis controlled by nuisance parameters. Since regression models
can have various forms, there is no general solution to this problem,
but it sometimes can be reduced to a so called ``reference
distribution'' by applying one of the three methods: (i) sampling from a
conditional distribution given a minimal sufficient statistic under
\(H_0\), (ii) parametric bootstrap sampling with nuisance parameters
estimated under \(H_0\), and (iii) Bayesian posterior predictive
sampling. The conditional distribution given a minimal sufficient
statistic is the best justified reference distribution among the three
\citep{buja_statistical_2009}. Essentially, null residuals can be
simulated by regressing \(N\) i.i.d standard normal random draws on the
predictors, then re-scaling it by the ratio of residual sum of square in
two regressions.

The effectiveness of lineup protocol for regression analysis has been
validated by \citet{majumder_validation_2013} under relatively simple
settings with up to two predictors. Their results suggest that visual
tests are capable of testing the significance of a single predictor with
a similar power to a t-test, though they express that in general it is
unnecessary to use visual inference if there exists a conventional test,
and they do not expect the visual test to perform equally well as the
conventional test. In their third experiment, where there is not a
conventional test, visual test outperforms the conventional test by a
large margin. This is encouraging, as it supports the use of visual
inference in situations where there are no existing statistical testing
procedures. Visual inference has also been integrated into diagnostics
for hierarchical linear models where the lineup protocol is used to
judge the assumptions of linearity, normality and constant error
variance for both the level-1 and level-2 residuals
\citep[\citet{loy2014hlmdiag} and
\citet{loy2015you}]{loy2013diagnostic}.

@@

The approach used in \citet{majumder_validation_2013} is as follows.
Define \(X_j = \{0,1\}\) to be a Bernoulli random variable measuring
whether participant \(j\) detected the data plot, and
\(X = \sum_{j=1}^{K}X_j\) be the total number of observers who detected
the data plot. Then, by imposing a relatively strong assumption that all
\(K\) evaluations are fully independent, under \(H_0\)
\(X \sim \mathrm{Binom}_{K,1/m}\). Therefore, the \(p\)-value of a
lineup of size \(m\) evaluated by \(K\) observer is estimated with
\(P(X \geq x) = 1 - F(x) + f(x)\), where \(F(.)\) is the binomial
cumulative distribution function, \(f(.)\) is the binomial probability
mass function and \(x\) is the realization of number of observers
choosing the data plot.

@@

The power of a visual test on the other hand, may additionally depend on
the ability of the particular participant, as the skill of each
individual may affect the number of observers who identify the data plot
from the lineup \citep{majumder_validation_2013}. To address this issue,
\citet{majumder_validation_2013} models the probability of participant
\(j\) correctly picking the data plot from lineup \(l\) using a
mixed-effect logistic regression, with participants treated as random
effect. Then, the estimated power of a visual test evaluated by a single
participant is the predicted value obtained from the mixed effects
model. However, this mixed effects model does not work with scenario
where participants are asked to select one or more most different plots.
In this scenario, having the probability of a participant \(j\)
correctly picking the data plot from a lineup \(l\) is insufficient to
determine the power of a visual test because it does not provide
information about the number of selections made by the participant for
the calculation of the \(p\)-value (See Equation
\ref{eq:pvalue-beta-binomial-approx}). Therefore, we directly estimate
the probability of a lineup being rejected by assuming that individual
skill has negligible effect on the variation of the power. This
assumption is not necessary true, but it helps to simplify the model
structure, thereby obviating a costly large-scale experiment to estimate
complex covariance matrices. The same model given in Equation
\ref{eq:logistic-regression-1-1} is applied to model the power of a
visual test.

@@

In each lineup evaluation, the participant was asked to select one or
more plots that are most different from others, provide a reason for
their selections, and evaluate how different they think the selected
plots are from others. If there is no noticeable difference between
plots in a lineup, participants were permitted to select zero plots
without providing the reason. No participant was shown the same lineup
twice. Information about preferred pronouns, age group, education, and
previous experience in visual experiments were also collected. A
participant's submission was only included in the analysis if the data
plot is identified for at least one attention check.

@@

Data collection used a total of 1152 lineups, and resulted in a total of
7974 evaluations from 443 participants. Roughly half corresponded to the
two models, non-linearity and heteroskedasticiy, and the three
collection periods had similar numbers of evaluations. Each participant
received two of the 24 attention check lineups which were used to filter
results of participants who were clearly not making an honest effort
(only 11 of 454). To estimate \(\alpha\) for calculating statistical
significance (see Section \ref{sig}) there were 720 evaluations of 36
null lineups. Neither the attention checks nor null lineups were used in
the subsequent analysis. The de-identified data, \texttt{vi\_survey}, is
made available in the R package, \texttt{visage}.

@@

Overall, the results show that the conventional tests are more sensitive
than the visual test. The conventional tests do have higher power for
the patterns they are designed to detect, and are generally unable to
detect other patterns. The visual test doesn't require specifying the
pattern ahead of time, relying purely on whether the observed residual
plot is detectably different from ``good'' residual plots. They will
perform equally well regardless of the type of model defect. This aligns
with the advice of experts on residual analysis, who consider residual
plot analysis to be an indispensable tool for diagnosing model problems.
What we gain from using a visual test for this purpose is the removal of
any subjective arguments about whether a pattern is visible or not. The
lineup protocol provides the calibration for detecting patterns: if the
pattern in the data plot cannot be distinguished from the patterns in
good residual plots, then no discernible problem with the model exists.

@@

The order of the polynomial is a primary factor contributing to the
pattern produced by the non-linearity model. Figure
\ref{fig:poly-power-uniform-j} explores the relationship between
polynomial order and power of the tests. The conventional tests have
higher power for lower orders of Hermite polynomials, and the power
drops substantially for the ``triple-U'' shape. To understand why this
is, we return to the application of the RESET test, which requires a
parameter indicating degree of fitted values to test for, and the
recommendation is to generically use four \citep{ramsey_tests_1969}.
However, the ``triple-U'' shape is constructed from the Hermite
polynomials using power up to 18. If the RESET test had been applied
using a higher power of no less than six, the power curve of
``triple-U'' shape will be closer to other power curves. This
illustrates the sensitivity of the conventional test to the parameter
choice, and highlights a limitation: it helps to know the data
generating process to set the parameters for the test, which is
unrealistic in practice. However, we examined this in more detail (see
Appendix) and found that there is no harm in setting the parameter
higher than four on the tests' operation for lower order polynomial
shapes. Using a parameter value of six, instead of four, yields higher
power regardless of generating process, and is recommended.

@@

The article was created with R packages \texttt{rticles}
\citep{rticles}, \texttt{knitr} \citep{knitr} and \texttt{rmarkdown}
\citep{rmarkdown}. The project's Github repository
(\url{https://github.com/TengMCing/lineup_residual_diagnostics})
contains all materials required to reproduce this article.

@@

\bibliographystyle{tfcad}
\bibliography{paper.bib}

>>>>

In addition, scatterplots of residuals plotted against the fitted values
and each of the explanatory variables, can be used to scrutinize their
relationships. If there are any visually discoverable patterns, the
model is potentially inadequate or incorrectly specified. In general,
one looks for noticeable departures from the model such as non-linear
pattern or heteroskedasticity. A non-linear pattern would suggest that
the model needs to have some additional non-linear terms.
Heteroskedasticity suggests that the error is dependent on the
predictors, and hence violates the independence assumption. However,
correctly judging whether NO pattern exists in a residual plot is a
difficult task for humans. Humans will almost always will see a pattern
(see \citet{kahneman}), so the question that really needs answering is
whether any pattern perceived is consistent with randomness, purely
sampling variability, or noise. It is especially difficult to teach this
to new analysts and students \citep{loy2021bringing}. To address this,
statistical tests are available to check for non-linear patterns
\citep[e.g.][]{ramsey1969tests}, and heteroskedasticity
\citep[e.g.][]{breusch1979simple}.

Linear regression is a well-established procedure, and there is
considerable literature describing diagnostic procedures, e.g.
\citet{draper1998applied}, \citet{montgomery1982introduction},
\citet{belsley1980regression}, \citet{cook1999applied} and
\citet{cook1982residuals}. Interestingly, despite the abundance of
conventional tests, ALL of these writings advise that plotting residuals
is an essential tool for diagnosing regression model problems:

@@

Graphical summaries where residuals are plotted against fitted values,
or other functions of the predictors (expected to be approximately
orthogonal to the residuals) are considered to be the most important
residual plots by \citet{cook1999applied}. Figure
\ref{fig:residual-plot-common-departures}A shows an example of an ideal
residual plot where points are symmetrically distributed around the
horizontal zero line (red), with no discernible patterns. There can be
various types of departures from this ideal pattern. Non-linearity,
heteroskedasticity and non-normality, shown in Figure
\ref{fig:residual-plot-common-departures}B,C,D are three commonly
checked departures.

@@

Figure \ref{fig:residual-plot-common-departures}D shows a scatterplot
where the residuals have a skewed distribution, as seen by the uneven
vertical spread. Unlike non-linearity and heteroskedasticity,
non-normality is usually detected with a different type of residual
plot: a histogram or normal probability plot. Because we focus on
scatterplots, non-normality is not one of the departures examined in
depth in this paper. (\citet{loy2016variations} discuss related work on
non-normality checking.)

@@

Many different hypothesis tests are available to detect specific model
defects. For example, the presence of heteroskedasticity can usually be
tested by applying the White test \citep{white1980heteroskedasticity} or
the Breusch-Pagan (BP) test \citep{breusch1979simple}, which are both
derived from the Lagrange multiplier test \citep{silvey1959lagrangian}
principle that relies on the asymptotic properties of the null
distribution. To test specific forms of non-linearity, one may apply the
F-test as a model structural test to examine the significance of
specific polynomial and non-linear forms of the predictors, or the
significance of proxy variables as in the Ramsey Regression Equation
Specification Error Test (RESET) \citep{ramsey1969tests}. The
Shapiro-Wilk (SW) normality test \citep{shapiro1965analysis} is the most
widely used test of non-normality included by many of the statistical
software programs. The Jarque-Bera test \citep{jarque1980efficient} is
also used to directly check whether the sample skewness and kurtosis
match a normal distribution.

Table \ref{tab:example-residual-plot-table} displays the \(p\)-values
from the RESET, BP and SW tests applied to the residual plots in Figure
\ref{fig:residual-plot-common-departures}. The RESET test and BP test
were computed using the \texttt{resettest} and \texttt{bptest} functions
from the R package \texttt{lmtest}, respectively. The SW test was
computed using the \texttt{shapiro.test} from the core R package
\texttt{stats}. \footnote{Although we didn't use it, it's useful to know
  that the R package \texttt{skedastic} \citep{skedastic} also contains
  a large collection of functions to test for heteroskedasticity.}
Although, the RESET test is exact, it requires the selection of a power
parameter. \citet{ramsey1969tests} recommends a power of four, which we
adopted in our analysis. The BP and SW tests are approximate.

We would expect the RESET test for non-linearity to reject residual plot
B, the BP test for heteroskedasticity to reject the residual plot C, and
the SW test for non-normality to reject residual plot D, which they all
do and all tests also correctly fail to reject residual plot A.
Interestingly, the BP and SW tests also reject the residual plots
exhibiting structure that they weren't designed for.
\citet{cook1982residuals} explain that most residual-based tests for
particular types of departure from model assumptions are also sensitive
to other types of departures. This could be considered a Type III error
\citep{kimball1957errors}, where the null hypothesis of good residuals
is correctly rejected but for the wrong reason. Also, some types of
departure can have elements of other types of departure, for example,
non-linearity could be viewed as heteroskedasticity. Additionally, other
data problems such as outliers can trigger rejection (or not) of the
null hypothesis \citep{cook1999applied}.

@@

The examination of data plots to infer signal or patterns (or lack
thereof) is fraught with variation in the human ability to interpret and
decode the information embedded in a graph
\citep{cleveland1984graphical}. Human examination of diagnostic plots
can feel subjective. Ideally we would expect visual evaluation to be
done objectively with small amounts of human-to-human variation in
reading data plots, but we cannot assume this is the case.

In practice, over-interpretation of a single plot is common. For
instance, \citet{roy2015using} described a published example where
authors over-interpreted separation between gene groups from a
two-dimensional projection of a linear discriminant analysis.
\citet{roy2015using} showed that there were no differences in the
expression levels between the gene groups. One solution to
over-interpretation is to examine the plot in the context of natural
sampling variability assumed by the model, called the lineup protocol,
as proposed in \citet{buja2009statistical}.
\citet{majumder2013validation} showed that the lineup protocol is
analogous to the null hypothesis significance testing framework. The
protocol consists of \(m\) randomly placed plots, where one plot is the
data plot, and the remaining \(m - 1\) plots, referred to as the
\emph{null plots}, are constructed using the same graphical procedure as
the data plot but the data is replaced with null data that is generated
in a manner consistent with the null hypothesis, \(H_0\). Then, an
observer who has not seen the data plot is asked to point out the most
different plot from the lineup. Under \(H_0\), it is expected that the
data plot would have no distinguishable difference from the null plots,
and the probability that the observer correctly picks the data plot is
\(1/m\). If one rejects \(H_0\) as the observer correctly picks the data
plot, then the Type I error of this test is \(1/m\). This protocol
requires a priori specification of \(H_0\) (or at least a null data
generating mechanism), much like the requirement of knowing the sampling
distribution of the test statistic in null hypothesis significance
testing framework.

@@

Data used in the \(m - 1\) null plots needs to be simulated. In
regression diagnostics, sampling data consistent with \(H_0\) is
equivalent to sampling data from the assumed model. As
\citet{buja2009statistical} suggested, \(H_0\) is usually a composite
hypothesis controlled by nuisance parameters. Since regression models
can have various forms, there is no general solution to this problem,
but it sometimes can be reduced to a so called ``reference
distribution'' by applying one of the three methods: (i) sampling from a
conditional distribution given a minimal sufficient statistic under
\(H_0\), (ii) parametric bootstrap sampling with nuisance parameters
estimated under \(H_0\), and (iii) Bayesian posterior predictive
sampling. The conditional distribution given a minimal sufficient
statistic is the best justified reference distribution among the three
\citep{buja2009statistical}. Essentially, null residuals can be
simulated by regressing \(N\) i.i.d standard normal random draws on the
predictors, then re-scaling it by the ratio of residual sum of square in
two regressions.

The effectiveness of lineup protocol for regression analysis has been
validated by \citet{majumder2013validation} under relatively simple
settings with up to two predictors. Their results suggest that visual
tests are capable of testing the significance of a single predictor with
a similar power to a t-test, though they express that in general it is
unnecessary to use visual inference if there exists a conventional test,
and they do not expect the visual test to perform equally well as the
conventional test. In their third experiment, where there is not a
conventional test, visual test outperforms the conventional test by a
large margin. This is encouraging, as it supports the use of visual
inference in situations where there are no existing statistical testing
procedures. Visual inference has also been integrated into diagnostics
for hierarchical linear models where the lineup protocol is used to
judge the assumptions of linearity, normality and constant error
variance for both the level-1 and level-2 residuals
\citep[\citet{loy2014hlmdiag} and
\citet{loy2015you}]{loy2013diagnostic}.

@@

The approach used in \citet{majumder2013validation} is as follows.
Define \(X_j = \{0,1\}\) to be a Bernoulli random variable measuring
whether participant \(j\) detected the data plot, and
\(X = \sum_{j=1}^{K}X_j\) be the total number of observers who detected
the data plot. Then, by imposing a relatively strong assumption that all
\(K\) evaluations are fully independent, under \(H_0\)
\(X \sim \mathrm{Binom}_{K,1/m}\). Therefore, the \(p\)-value of a
lineup of size \(m\) evaluated by \(K\) observer is estimated with
\(P(X \geq x) = 1 - F(x) + f(x)\), where \(F(.)\) is the binomial
cumulative distribution function, \(f(.)\) is the binomial probability
mass function and \(x\) is the realization of number of observers
choosing the data plot.

@@

The power of a visual test on the other hand, may additionally depend on
the ability of the particular participant, as the skill of each
individual may affect the number of observers who identify the data plot
from the lineup \citep{majumder2013validation}. To address this issue,
\citet{majumder2013validation} models the probability of participant
\(j\) correctly picking the data plot from lineup \(l\) using a
mixed-effect logistic regression, with participants treated as random
effect. Then, the estimated power of a visual test evaluated by a single
participant is the predicted value obtained from the mixed effects
model. However, this mixed effects model does not work with scenario
where participants are asked to select one or more most different plots.
In this scenario, having the probability of a participant \(j\)
correctly picking the data plot from a lineup \(l\) is insufficient to
determine the power of a visual test because it does not provide
information about the number of selections made by the participant for
the calculation of the \(p\)-value (See Equation
\ref{eq:pvalue-beta-binomial-approx}). Therefore, we directly estimate
the probability of a lineup being rejected by assuming that individual
skill has negligible effect on the variation of the power. This
assumption essentially averages out the subject ability and helps to
simplify the model structure, thereby obviating a costly large-scale
experiment to estimate complex covariance matrices. The same model given
in Equation \ref{eq:logistic-regression-1-1} is applied to model the
power of a visual test.

@@

In each lineup evaluation, the participant was asked to select one or
more plots that are most different from others, provide a reason for
their selections, and evaluate how different they think the selected
plots are from others. If there is no noticeable difference between
plots in a lineup, participants had the option to select zero plots
without the need to provide a reason. During the process of recording
the responses, a zero selection was considered to be equivalent to
selecting all 20 plots. No participant was shown the same lineup twice.
Information about preferred pronouns, age group, education, and previous
experience in visual experiments were also collected. A participant's
submission was only included in the analysis if the data plot is
identified for at least one attention check.

@@

Data collection used a total of 1152 lineups, and resulted in a total of
7974 evaluations from 443 participants. Roughly half corresponded to the
two models, non-linearity and heteroskedasticiy, and the three
collection periods had similar numbers of evaluations. Each participant
received two of the 24 attention check lineups which were used to filter
results of participants who were clearly not making an honest effort
(only 11 of 454). To estimate \(\alpha\) for calculating statistical
significance (see Section \ref{sig}) there were 720 evaluations of 36
null lineups. Neither the attention checks nor null lineups were used in
the subsequent analysis. The de-identified data, \texttt{vi\_survey}, is
made available in the R package, \texttt{visage}.\footnote{The package
  \texttt{visag} is currently available on GitHub
  (\url{https://github.com/TengMCing/visage})}

@@

Overall, the results show that the conventional tests are more sensitive
than the visual test. The conventional tests do have higher power for
the patterns they are designed to detect, but they typically fail to
detect other patterns unless those patterns are particularly strong. The
visual test doesn't require specifying the pattern ahead of time,
relying purely on whether the observed residual plot is detectably
different from ``good'' residual plots. They will perform equally well
regardless of the type of model defect. This aligns with the advice of
experts on residual analysis, who consider residual plot analysis to be
an indispensable tool for diagnosing model problems. What we gain from
using a visual test for this purpose is the removal of any subjective
arguments about whether a pattern is visible or not. The lineup protocol
provides the calibration for detecting patterns: if the pattern in the
data plot cannot be distinguished from the patterns in good residual
plots, then no discernible problem with the model exists.

@@

The order of the polynomial is a primary factor contributing to the
pattern produced by the non-linearity model. Figure
\ref{fig:poly-power-uniform-j} explores the relationship between
polynomial order and power of the tests. The conventional tests have
higher power for lower orders of Hermite polynomials, and the power
drops substantially for the ``triple-U'' shape. To understand why this
is, we return to the application of the RESET test, which requires a
parameter indicating degree of fitted values to test for, and the
recommendation is to generically use four \citep{ramsey1969tests}.
However, the ``triple-U'' shape is constructed from the Hermite
polynomials using power up to 18. If the RESET test had been applied
using a higher power of no less than six, the power curve of
``triple-U'' shape will be closer to other power curves. This
illustrates the sensitivity of the conventional test to the parameter
choice, and highlights a limitation: it helps to know the data
generating process to set the parameters for the test, which is
unrealistic in practice. However, we examined this in more detail (see
Appendix) and found that there is no harm in setting the parameter
higher than four on the tests' operation for lower order polynomial
shapes. Using a parameter value of six, instead of four, yields higher
power regardless of generating process, and is recommended.

@@

The article was created with R packages \texttt{rticles}
\citep{rticles}, \texttt{knitr} \citep{knitr} and \texttt{rmarkdown}
\citep{rmarkdown}. The project's GitHub repository
(\url{https://github.com/TengMCing/lineup_residual_diagnostics})
contains all materials required to reproduce this article.

@@

\bibliographystyle{tfcad}
\bibliography{ref.bib}
